A Logistic Regression Approach to CoIL Challenge 2000

### 2019-12-09 final project/Draft.Rmd
Updated Draft.RMD
Added some correlation to the Draft

---
title             : "A Logistic Regression Approach to CoIL Challenge 2000"
shorttitle        : " "
author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
abstract: |
  A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
  <!-- https://tinyurl.com/ybremelq 
    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  -->
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  word_document:
    reference_docx: word-template.docx
  papaja::apa6_pdf:
    #citation_package: natbib #biblatex
  html_document: default
---

@@

df <- df %>%
  mutate(MOSTYPE = as.factor(MOSTYPE),
         MGEMLEEF = as.factor(MGEMLEEF),
         MOSHOOFD = as.factor(MOSHOOFD),
         MGODRK = as.factor(MGODRK),
         PWAPART = as.factor(PWAPART),
         CARAVAN = as.factor(CARAVAN)) %>%
  mutate(MOSTYPE = recode(MOSTYPE, !!!MOSTYPE_labels),
         MGEMLEEF = recode(MGEMLEEF, !!!MGEMLEEF_labels),
         MOSHOOFD = recode(MOSHOOFD, !!!MOSHOOFD_labels),
         MGODRK = recode(MGODRK, !!!MGODRK_labels),
         PWAPART = recode(PWAPART, !!!PWAPART_labels))
}

@@

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
                                          104, 105, 106, 107, 108, 109, 110,
                                          111, 112, 115, 121),
                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
                                      1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122)
```

@@

@@

The dataset is unbalanced.  There are only `r df %>% filter(CARAVAN == 1) %>% nrow()` cases that have purchased the mobile home insurance policy and `r df %>% filter(CARAVAN == 0) %>% nrow()` that have not.

@@

for (var in categorical_vars){
  df$var <- df[,var]
  p <- ggplot(df) +
    geom_mosaic(aes(x = product(CARAVAN, var), fill=CARAVAN), na.rm=TRUE) +
    labs(x=var, y="CARAVAN") +
    theme(legend.position = "none") + 
    scale_fill_brewer(palette = "Set1")

@@

@@

@@

Most of the correlation heatmap is yellow which is has a correlation coeffient of zero.  Here's a density plot of the correlation coefficients:

@@

>>>>

---
title: "A Logistic Regression Approach to CoIL Challenge 2000"
author: "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation: MS in Data Science Students CUNY School of Professional Studies
output:
  word_document:
    reference_docx: word-template.docx
  pdf_document: default
  papaja::apa6_pdf: null
  html_document: default
classoption: man
documentclass: apa6
draft: no
figurelist: no
floatsintext: no
footnotelist: no
keywords: CoIL Challenge, Logistic Regression
linenumbers: no
mask: no
bibliography: Manuscript.bib
shorttitle: DRAFT SUBMISSION
tablelist: no
abstract: "A logistic regression based solution to the CoIL Challenge 2000 is described.
  The challenge consists of correctly identifying potential customers for an insurance
  \n  One sentence summarizing the main result (with the words \"**here we show**\"
  or their equivalent).\n\nTwo or three sentences explaining what the **main result**
  reveals in direct comparison to what was thought to be the case previously, or how
  the  main result adds to previous knowledge.\n\nOne or two sentences to put the
  results into a more **general context**.\n\nTwo or three sentences to provide a
  **broader perspective**, readily comprehensible to a scientist in any discipline.\n-->\n"
  product, and describing their characteristics.\n<!-- https://tinyurl.com/ybremelq
wordcount: X
---

@@

df <- df %>%
  mutate(MOSTYPE = as.factor(MOSTYPE),
         MGEMLEEF = as.factor(MGEMLEEF),
         MOSHOOFD = as.factor(MOSHOOFD),
         MGODRK = as.factor(MGODRK),
         PWAPART = as.factor(PWAPART),
         CARAVAN = as.factor(CARAVAN)) %>%
  mutate(MOSTYPE = recode_factor(MOSTYPE, !!!MOSTYPE_labels),
         MGEMLEEF = recode_factor(MGEMLEEF, !!!MGEMLEEF_labels),
         MOSHOOFD = recode_factor(MOSHOOFD, !!!MOSHOOFD_labels),
         MGODRK = recode_factor(MGODRK, !!!MGODRK_labels),
         PWAPART = recode_factor(PWAPART, !!!PWAPART_labels))
}

@@

```{r, echo=FALSE}
# Recreate the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83,
                                          86, 91, 94, 95, 96, 97, 99, 101, 103,
                                          104, 105, 106, 107, 108, 109, 110,
                                          111, 112, 115, 121),
                                n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3,
                                      1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1,
                                      1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity") +
  xlab("Number of Correct Predictions") +
  ylab("CoIL Challenge Submissions") +
  xlim(37, 122) +
  labs(title="Frequency Distribution of Prediction Scores")
```

@@

Our response variable here is the `CARAVAN` variable. It represents whether the person in question purchased an caravan insurance plan.

@@

The dataset is unbalanced as there are only `r df %>% filter(CARAVAN == 1) %>% nrow()` cases that have purchased the mobile home insurance policy and `r df %>% filter(CARAVAN == 0) %>% nrow()` that have not.

@@

for (var in categorical_vars){
  df$var <- df[,var]

  p <- ggplot(df) +
    geom_mosaic(aes(x = product(CARAVAN,var), fill=CARAVAN),na.rm=TRUE) +
    labs(x=var, y="CARAVAN") +
    theme(legend.position = "none") + 
    scale_fill_brewer(palette = "Set1") +
    theme(axis.text.x = element_text(angle=45,hjust=1))

@@

Looking at our categorical variables, there is some variation in frequency of `CARAVAN` in some values of `MOSTYPE` and `MGEMLEEF`, but the rest have quite a bit less variation.

@@

For numeric variables, we don't see any which can clearly seperate the `CARAVAN` variable values.

### Correlation

@@

Most of the correlation heatmap is yellow which is has a correlation coeffient of zero. There are a few variable combinations with correlation, however. Here's a density plot of the correlation coefficients:

@@

Next we will look at variable correlation with the outcome variable:

corrs <- data.frame(CARAVAN = 1)

for (var in numeric_vars){
   t <- df %>% select(CARAVAN,var) %>% mutate(CARAVAN = as.numeric(CARAVAN))
   corrs[,var] <- cor(t,use="complete.obs")[1,2]
}

corrs <- corrs %>% gather(key="variable",value="correlation",-CARAVAN) %>%
  select(-CARAVAN) %>% mutate(positive = ifelse(correlation > 0,TRUE,FALSE))

corrs[1:20,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
  geom_col() + scale_y_continuous(limits=c(-1,1)) +
  scale_fill_manual(values=c("#d13a26","#369970")) +
  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
  labs(title="Variable Correlations with Dependent Variable",
       subtitle="Part 1")

corrs[21:40,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
  geom_col() + scale_y_continuous(limits=c(-1,1)) +
  scale_fill_manual(values=c("#d13a26","#369970")) +
  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
  labs(title="Variable Correlations with Dependent Variable",
       subtitle="Part 2")

corrs[41:60,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
  geom_col() + scale_y_continuous(limits=c(-1,1)) +
  scale_fill_manual(values=c("#d13a26","#369970")) +
  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
  labs(title="Variable Correlations with Dependent Variable",
       subtitle="Part 3")

corrs[61:80,] %>% ggplot(aes(x=variable,y=correlation,fill=positive)) +
  geom_col() + scale_y_continuous(limits=c(-1,1)) +
  scale_fill_manual(values=c("#d13a26","#369970")) +
  theme(legend.position = "None", axis.text.x = element_text(angle=90)) +
  labs(title="Variable Correlations with Dependent Variable",
       subtitle="Part 4")

```

None of the numeric variables, alone, seem to be well-correlated with the dependent variable. In fact, the correlation is quite poor across the board:

```{r, echo = FALSE}
summary(corrs$correlation)
```
